---
title: "Logging"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Logging}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  results = 'markup'
)
```

Logging is a very important part of a server application - it allows to analyze workload, collect metrics and facilitates finding critical bugs. 

`RestRserve` ships with basic (yet handy) logging utility - `Logger`. It can be used in the application code, but what is more important, it is integrated into `Application` class in order to print useful messages during app execution.


Each applications comes with built-in logger member:

```{r}
library(RestRserve)
application = Application$new()
application$logger$info("hello from logger")
```
`logger` is an instance of `Logger` class. We believe it is important to have logs in a machine readable format, so by default logger writes messages to `stdout` in JSON format. 

Amount of logging can be adjusted with `set_log_level` method:
```{r}
application$logger$set_log_level("error")
application$logger$info("you won't see this message")
```

If you want to customize `logger`, you can tweak it with `logger$set_printer()` function or even swap it to your own (see `?Logger` for details):

```{r}
application$logger = Logger$new(level = "trace", name = "mylogger")
```
Moreover `Logger` mimics loggers from fully-featured [lgr](https://cran.r-project.org/package=lgr) package - you can use them interchangeably.

`application$logger` object however is not designed to be called by end user to produce messages. It is used internally in the `application` to emit status messages along the request-response processing pipeline.

Consider following example. We develop a method which takes a square root of a number:
```{r}
application = Application$new()
application$add_get("/sqrt", function(request, response) {
  response$body = sqrt(x)
})
```
Now let's query it:
```{r}
# let's emulate query string "/sqrt?x=10"
request = Request$new(path = "/sqrt", method = "GET", parameters_query = list(x = "10")) 
response = application$process_request(request)
response$body
```
Surprise! (or not?)

All request query parameters are strings, but we forgot to convert `x` to a numeric value. This caused R interpreter to throw an error. 

Note however how gracefully RestRserve handled this case:

- it hasn't crashed and produced HTTP error with 500 code
- written error with expanded traceback to the log. This is particularly important as R's traceback does not contain the full stack trace if you catch errors and warnings!

For example following code will just give you last error (but not a traceback):
```{r}
fun2 = function(x) {
  sqrt(x)
}
fun1 = function(x) {
  fun2(x)
}
try(fun1('a'))
```
Which is not very useful when you try to debug server application.


In contrast let's call it within RestRserve application:
```{r}
application$add_get("/sqrt2", function(request, response) {
  response$body = fun1(x)
})
request = Request$new(path = "/sqrt2", method = "GET", parameters_query = list(x = "10")) 
response = application$process_request(request)
```

Now we clearly see that error happens inside `fun2`
