---
title: "Serving machine learning models"
author: "Dmitriy Selivanov"
date: "`r Sys.Date()`"
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

This example is a good starting point for all the developers who want to understand how RestRserve works. All the code can be found in the [example directory](https://github.com/dselivanov/RestRserve/tree/master/examples/xgboost-mushrooms) of the RestRserve repository.

In many cases the end goal of data scientists / data engineers is to create a *data product* which will be used by other people or services/applications. `RestRserve` focuses on the later.


# Pipeline

Here we show a simplified pipeline for creating a microservice which will serve a machine learning model.

Assume we need to be able to predict something based on the data. First we need to build a model:

1. Get raw data
1. Perform set of transforamtions - extract features, encode variables, etc.
1. Train a model on a transformed data

Than after several iterations of the steps above we may have a good enough model. Now we want it process new data. This could be decomposed into:

1. Get raw data
1. Perform **exactly the same transforamtions** as we've done during the training
1. Apply pre-trained model

As we can see we have to think about reproducibility of the data transformations and about how to save/load the model. Fortunately almost all R objects can be serialized and stored to disk with `saveRDS()` function. After we can read them with `readRDS()` function. There are some important remarks here:

- need to pay attention to **closures** - need to make sure that all necessary data/code is captured during the serialization
- **not all the R objects are serializable**! Good examples are connections and external pointers.

# Example

We will use [xgboost](https://github.com/dmlc/xgboost) library here to demonstrate how to build RestRserve application. Also we will use [Mushroom Data Set](https://archive.ics.uci.edu/ml/datasets/Mushroom) which is built-in into R's `xgboost` package.

Given the description of the mushrooms in terms of physical characteristics the task is to predict whether mushrom is poisonous or edible. 

## Data

Data in the package is stored as a CSC sparse matrix:

```{r, eval=T}
library(xgboost)
library(Matrix)
data(agaricus.train, package = "xgboost")
str(agaricus.train$data)
```

All the mushroom properties are "one-hot-encoded":
```{r, eval=T}
features = colnames(agaricus.train$data)
head(features)
```

## Train the model

Let's train our model (for simplicity we omit hyper-parameter selection / cross-validation step):

```{r}
bst = xgboost(data = agaricus.train$data, 
              label = agaricus.train$label, 
              max_depth = 2, eta = 1,
              nrounds = 5, 
              objective = "binary:logistic")
```

Now we need to save the model in order to be able to use it during inference time. `xgboost` models can't be serialized with `saveRDS()`, but fortunately library provides its own save and load methods:

```{r}
xgb.save(bst, "model.xgb")
```

## Query the service

Now we want to wrap model a http-microservice. We want to send a http-request with the properties of the mushroom and service should respond with a probability of the mushroom being edible. Properties can be send to the server using different http methods (usually GET or POST). Here we will demonstrate

- POST method with the payload in JSON format. 
- GET method with parameters encoded inside the query string

POST request can look like:

`'[{"featureName": "stalk-color-below-ring", "featureValue": "red"},{"featureName": "cap-shape","featureValue": "conical"}]'`

GET parameter can look like:

`stalk-color-below-ring=red;cap-shape=conical`. 

Note that in reality `;` and `=` symbols need to be url-encoded. So query parameter will look like:

`stalk-color-below-ring%3Dred%3Bcap-shape%3Dconical`.

In order to be able to make prediction we need to transform input raw data to the format that our model can consume - CSC sparse matrix:

```{r}
# this function create a function which can be used to 
# encode input json to a sparse matrix
model_matrix_constructor = function(features) {
  n_features = length(features)
  function(x) {
    # read key-value pairs from json
    key_val = jsonlite::fromJSON(x, simplifyVector = FALSE)
    # make them  "key=value" string as it was in original feature space
    for(i in length(key_val)) {
      kv = key_val[[i]]
      key_val[[i]] = paste(kv$key, kv$val, sep = "=")
    }
    # find the column index for a given feature
    col_index = match(key_val, features)
    # if we've received unseen priperties - ignore them 
    col_index = col_index[!is.na(col_index)]
    # since request contains information about single observation -
    # there should be 1 row in the matrix
    row_index = rep(1, length(col_index))
    # cosntruct sparse matrix as it is required by xgboost
    sparseMatrix(i = row_index, 
                 j = col_index, 
                 x = 1, 
                 dims = c(1, n_features))

  }
}

create_model_matrix = model_matrix_constructor(features)
```

Save transformation function in order to use it during inference time:
```{r}
saveRDS(create_model_matrix, "create_model_matrix.rds")
```

## RestRserve application

Now everything is ready to build and start our web service! Logically what we will do:

1. load necessary libraries
1. load model and function which transforms raw data to the format which model can consume
1. create functions which will handle incoming requests
    - add annotations in YAML format to generate OpenAPI specification (used for swagger-ui)
1. register function to handle particular end points
1. generate OpenAPI specification (RestRserve can automatically parse function annotations) and create swagger-ui

```{r}
library(RestRserve)
library(xgboost)
library(Matrix)

create_model_matrix = readRDS("create_model_matrix.rds")
bst = xgb.load("model.xgb")

pred_xgb_post = function(request, response) {
  #' ---
  #' description: predict whether mushroom edible or poisonous
  #' operationId: postXgbPredict
  #' requestBody:
  #'   description: "mushroom properties"
  #'   required: true
  #'   content:
  #'     application/json:
  #'       schema:
  #'         type: array
  #'         items:
  #'           required:
  #'             - featureName
  #'             - featureValue
  #'           properties:
  #'             featureName:
  #'               type: string
  #'               example: "stalk-color-below-ring"
  #'             featureValue:
  #'               type: string
  #'               example: "red"
  #' responses:
  #'   200:
  #'     description: API response
  #'     content:
  #'       text/plain:
  #'         schema:
  #'           type: numeric
  #'           example: 0.5
  #' ---
  x = rawToChar(request$body)
  x = create_model_matrix(x)
  response$body = as.character(predict(bst, x))
  response$content_type = "text/plain"
  response$headers = character(0)
  response$status_code = 200L
}
pred_xgb_get = function(request, response) {
  #' ---
  #' description: predict whether mushroom edible or poisonous
  #' operationId: getXgbPredict
  #' parameters:
  #'   - name: "mushrom-properties"
  #'     description: >
  #'       String of the 'key=value' pairs of mushroom properties.
  #'       Several paramenters can be provided - use semicolon ';'
  #'       to concatenate them
  #'     in: query
  #'     schema:
  #'       type: string
  #'     example: "stalk-color-below-ring=red;cap-shape=conical"
  #' responses:
  #'   200:
  #'     description: API response
  #'     content:
  #'       text/plain:
  #'         schema:
  #'           type: numeric
  #'           example: 0.5
  #' ---

  # create json in order to be able to re-use create_model_matrix() function
  # just for demonstration - by far not ideal solution!
  x = strsplit(request$query[["mushrom-properties"]], ";", fixed = TRUE)[[1]]
  x = strsplit(x, "=", fixed = TRUE)
  x = jsonlite::toJSON (lapply(x, function(z) list(featureName = z[[1]], featureValue = z[[2]])), auto_unbox = TRUE)
  # create feature matrix
  x = create_model_matrix(x)
  response$body = as.character(predict(bst, x))
  response$content_type = "text/plain"
  response$headers = character(0)
  response$status_code = 200L
}
#------------------------------------------------------------------------------------------
# create application
#------------------------------------------------------------------------------------------
RestRserveApp = RestRserve::RestRserveApplication$new()
RestRserveApp$add_post(path = "/predict", FUN = pred_xgb_post)
RestRserveApp$add_get(path = "/predict", FUN = pred_xgb_get)
RestRserveApp$add_openapi()
RestRserveApp$add_swagger_ui("/")
RestRserveApp$run(8001)
```

# Try it out

Now we can visit [http://localhost:8001/](http://localhost:8001/) and try our application.

# Stress test

Also we can test on how many requests per second it can handle (we use [apib](https://github.com/apigee/apib)):

```{bash, eval=FALSE}
echo '[{"featureName": "stalk-color-below-ring", "featureValue": "red"},{"featureName": "cap-shape","featureValue": "conical"}]' > request.json

apib --method POST --input-file request.json -c 16 -d 5 http://localhost:8001/predict

rm request.json
```

And stress-test GET method:
```{bash, eval=FALSE}
apib -c 16 -d 5 http://localhost:8001/predict?mushrom-properties=stalk-color-below-ring%3Dred%3Bcap-shape%3Dconical
```
